Complete the following questions for Assignment 2. Please upload your completed Assignment to GitHub classrooms as **R MarkDown Files**.

1. Create a vector called score that contains every 9th number between 1 and 109.
```{r}
seq(1,109, by=9)
score<- c(10,19,28,37,46,55,64,73,82,91,100,109)
```
score<- c(seq(1,109, by=9)) works but it gives 13 rows not 12, so I get errors. I'm assuming it's not supposed to include 1?
2. Create a vector called participants that contains the numbers 1-12.
```{r}
participants<- c(seq(1,12))
participants

```
3. Create a vector called condition that repeats the words control and experimental six times. 
```{r}
condition<- c(rep("control", 6), rep("experimental", 6))
condition
```
4. Create a dataframe from the vectors in #1-3. 
```{r}
dat<- cbind.data.frame(score,participants,condition)
dat
```
5. Take the mean of the score column.
```{r}
mean(score)
59.5
```
6. Open the mtcars dataframe. Create a subset of data for cars that have 6 cylinders (cyl column).
```{r}
data(mtcars)
mtcars[mtcars$cyl==6,]
cylSix<- c(subset(mtcars, mtcars$cyl==6))
```
7. Summarize this subset of data. 
```{r}
summary(cylSix)
```
8. Create a dataframe that contains cars with higher than average miles per gallon (mpg). 
```{r}
mean(mtcars$mpg)
data(mtcars)
mpgHigh<- mtcars[mtcars$mpg>20.09062,]
mpgHigh
```
9. How many cars are in this subset? 
14

Challenge Questions: 
10. Open the diamonds dataset. Create a new dataframe that has diamonds greater than 1 carat and an ideal cut. Find the mean and standard deviation of price for this data. 
```{r}
data(diamonds)
```
11. Create a new dataframe that includes all diamonds with a price between $2500 and $5000. Which color occurs most in this subset? 